---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE
)
```
# opendatatoronto

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/sharlagelfand/opendatatoronto.svg?branch=master)](https://travis-ci.org/sharlagelfand/opendatatoronto)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/sharlagelfand/opendatatoronto?branch=master&svg=true)](https://ci.appveyor.com/project/sharlagelfand/opendatatoronto)
[![Codecov test coverage](https://codecov.io/gh/sharlagelfand/opendatatoronto/branch/master/graph/badge.svg)](https://codecov.io/gh/sharlagelfand/opendatatoronto?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/opendatatoronto)](https://cran.r-project.org/package=opendatatoronto)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

`opendatatoronto` is an R interface to the [City of Toronto Open Data Portal](https://portal0.cf.opendata.inter.sandbox-toronto.ca/). The goal of the package is to help read data directly into R without needing to manually download it via the portal.

## Installation

You can install the development version from GitHub with:

``` r
# install.packages("remotes")
remotes::install_github("sharlagelfand/opendatatoronto")
```

## Example

In the Portal, datasets are called **packages**. You can see a list of available packages by using `list_packages()`:

```{r list-packages-example}
library(opendatatoronto)
packages <- list_packages(limit = 100)
packages
```

You can also search packages by title:

```{r search-packages-example}
ttc_packages <- search_packages("ttc")

ttc_packages
```

Within a package, there are a number of **resources** - e.g. CSV, XSLX, JSON, SHP files, and more. Resources are the actual "data".

For a given package, you can get a list of resources using `list_package_resources()`:

```{r list-marriage-license-resources}
library(dplyr)

marriage_licence_packages <- packages %>%
  filter(title == "Marriage Licence Statistics")

marriage_licence_resources <- marriage_licence_packages %>%
  list_package_resources()

marriage_licence_resources
```

Finally (and most usefully!), you can download the resource (i.e., the actual data) directly into R using `get_resource()`:

```{r get-marriage-licenses}
marriage_licence_statistics <- marriage_licence_resources %>%
  get_resource()

marriage_licence_statistics
```
