---
title: "Retrieving multi-sheet XLS/XLSX resources"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Retrieving multi-sheet XLS/XLSX resources}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

Excel files (XLS and XLSX) are a common form of data on the [City of Toronto Open Data Portal](https://open.toronto.ca).

In cases where the file only contains one sheet, the resource is returned as a tibble. For example, this data set on [TTC Ridership Analysis from 1985 to 2018](https://open.toronto.ca/dataset/ttc-ridership-analysis/):

```{r setup}
library(opendatatoronto)

search_packages("TTC Ridership Analysis") %>%
  list_package_resources() %>%
  get_resource()
```

When the file contains multiple sheets, the resource is returned as a named list, where the names are the names of the sheets, as in the dataset on [Wellbeing Toronto Demographics](https://open.toronto.ca/dataset/wellbeing-toronto-demographics/):

```{r}
library(dplyr)

wellbeing_toronto_demographics <- search_packages("Wellbeing Toronto - Demographics") %>%
  filter(title == "Wellbeing Toronto - Demographics") %>%
  list_package_resources() %>%
  filter(name == "wellbeing-toronto-demographics") %>%
  get_resource()

str(wellbeing_toronto_demographics, max.level = 1)
```

To access the relevant sheet, pull out the list element:

```{r}
wellbeing_toronto_demographics[["IndicatorMetaData"]]
```

There are also cases where the file contains multiple sheets and it would be helpful to have them all together as a single data set. For example, the [2019 TTC Bus Delay Data](https://open.toronto.ca/dataset/ttc-bus-delay-data/):

```{r}
ttc_bus_delays_2019 <- search_packages("TTC Bus Delay Data") %>%
  list_package_resources() %>%
  filter(name == "ttc-bus-delay-data-2019") %>%
  get_resource()
```

The result of is a list with an element for every month of data, each of which is a tibble:

```{r}
str(ttc_bus_delays_2019, max.level = 1)
```

Note that the data for for the element `Apr 2019` has one more variable than the rest (11 versus 10):

```{r}
sapply(ttc_bus_delays_2019, colnames)
```

It seems that the `Apr 2019` data has gained a variable `Incident ID`, and that the variables `Min Gap` and `Min Delay`, present in all the other months, have been renamed to `Gap` and `Delay`, respectively.

We can rename these two variables:

```{r}
ttc_bus_delays_2019[["Apr 2019"]] <- ttc_bus_delays_2019[["Apr 2019"]] %>%
  rename(`Min Gap` = Gap, `Min Delay` = Delay)
```

and combine all of the elements into a single tibble using `dplyr::bind_rows()`:

```{r}
ttc_bus_delays_2019_combined <- bind_rows(ttc_bus_delays_2019)

ttc_bus_delays_2019_combined
```

Unfortunately, it looks like the `Time` variable got Excel™ed, and will need some data cleaning.

For interests sake, it appears that `Incident ID` is a lookup ID for the type of incident -- only present in the `Apr 2019` data, but interesting nonetheless!

```{r}
ttc_bus_delays_2019_combined %>%
  filter(!is.na(`Incident ID`)) %>%
  distinct(`Incident ID`, Incident)
```
