---
title: "Retrieving multiple resources using purrr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Retrieving multiple resources using purrr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

It may be useful to retrieve multiple or all of the resources for a package. For example, the [TTC Subway Delay Data](https://open.toronto.ca/dataset/ttc-subway-delay-data/) is released monthly, with a new file each month.

```{r setup}
library(opendatatoronto)

ttc_subway_delays_resources <- search_packages("TTC Subway Delay Data") %>%
  list_package_resources()

ttc_subway_delays_resources
```

It would be cumbersome to retrieve them one at a time, e.g. via 

```{r, eval = FALSE}
library(dplyr)

delays_jan_2019 <- ttc_subway_delays_resources %>%
  filter(name == "ttc-subway-delay-january-2019") %>%
  get_resource()

delays_feb_2019 <- ttc_subway_delays_resources %>%
  filter(name == "ttc-subway-delay-february-2019") %>%
  get_resource()
```

and so on.

Instead, you can iterate through the different resources using the [`purrr` package](https://purrr.tidyverse.org/) and get them in a nested tibble:

```{r include = FALSE}
library(dplyr)
```

```{r}
library(purrr)

ttc_subway_delays_2019 <- ttc_subway_delays_resources %>%
  filter(grepl("2019", name)) %>%
  mutate(data = map(id, get_resource))

ttc_subway_delays_2019
```

All of these resources contain equivalent data, with the same column names

```{r}
map(ttc_subway_delays_2019[["data"]], colnames)
```

so the data can easily be unnested and combined:

```{r}
bind_rows(ttc_subway_delays_2019[["data"]])
```

