---
title: "Introduction to opendatatoronto"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to opendatatoronto}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

`opendatatoronto` is a package for accessing the [City of Toronto Open Data portal](https://open.toronto.ca/) from within R.

In the portal, datasets are called **packages**. You can see a list of available packages by using `list_packages()`:

```{r list-packages-example}
library(opendatatoronto)

packages <- list_packages(limit = 50)

packages
```

Or, you can search packages by title using `search_packages()`:

```{r}
apartment_packages <- search_packages("Apartment")

apartment_packages
```

Within a package, there are a number of **resources** - e.g. CSV, XSLX, JSON, SHP files, and more. Resources are the actual "data".

For a given package, you can get a list of resources using `list_package_resources()`:

```{r list-apartment-building-registration-resources}
library(dplyr)

apartment_building_registration_package <- apartment_packages %>%
  filter(title == "Apartment Building Registration")

apartment_building_registration_resources <- apartment_building_registration_package %>%
  list_package_resources()

apartment_building_registration_resources
```

Finally (and most usefully!), you can download the resource (i.e., the actual data) directly into R using `get_resource()`:

```{r get-apartment-building-registration}
apartment_building_registration_data <- apartment_building_registration_resources %>%
  get_resource()

apartment_building_registration_data
```

The `opendatatoronto` package can currently handle the download of CSV, XLS/XLSX, XML, JSON, SHP, and GeoJSON resources, as well as ZIP resources that contain multiple files. For more information, see the following vignettes:

* [Multi-sheet XLS/XLSX resources](multisheet_resources.html)
* [Multi-file ZIP resources](multifile_zip_resources.html)
* [Spatial data](spatial_data.html)
* [Retrieving multiple resources using purrr](multiple_resources_purrr.html)
