---
title: "Working with spatial data from the portal"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with spatial data from the portal}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

There is a ton of spatial data on the [City of Toronto Open Data Portal](https://open.toronto.ca). Spatial resources are retrieved the same way as all other resources, by using `get_resource()`, and may require the `sf` and `geojsonsf` packages.

We can look at [bicycle parking](https://open.toronto.ca/dataset/bicycle-parking-racks/) in Toronto. The result is an `sf` object with WGS84 projection.

```{r setup}
library(opendatatoronto)
library(dplyr)

bike_parking_racks <- search_packages("Bicycle Parking Racks") %>%
  list_package_resources() %>%
  filter(name == "Bicycle Parking Racks Data") %>%
  get_resource()

bike_parking_racks
```

Using the [`cancensus` package](https://github.com/mountainMath/cancensus), we can plot the bike racks along with a map of Toronto!

```{r fig.width = 7}
library(ggplot2)
library(cancensus)

toronto <- get_census(
  dataset = "CA16", regions = list(CMA = "35535"),
  level = "CSD", geo_format = "sf", quiet = TRUE
)

ggplot() +
  geom_sf(data = toronto) +
  geom_sf(data = bike_parking_racks) +
  theme_minimal()
```
