---
title: "Introduction to opendatatoronto"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

`opendatatoronto` is an R interface to the [City of Toronto Open Data Portal](https://open.toronto.ca/). The goal of the package is to help read data directly into R without needing to manually download it via the portal. 

In the portal, datasets are called **packages**. You can see a list of available packages by using `list_packages()`. This will show metadata about the package, including what topics (i.e. tags) the package covers, a description of it,  any civic issues it addresses, how many resources there are (and their formats), how often it is is refreshed and when it was last refreshed.

```{r list-packages-example}
library(opendatatoronto)

packages <- list_packages(limit = 10)

packages
```

Or, you can search packages by title using `search_packages()`:

```{r}
apartment_packages <- search_packages("Apartment")

apartment_packages
```

You can also see metadata for one specific package using `show_package()`:

```{r}
show_package("996cfe8d-fb35-40ce-b569-698d51fc683b")
```

Within a package, there are a number of **resources** - e.g. CSV, XSLX, JSON, SHP files, and more. Resources are the actual "data".

For a given package, you can get a list of resources using `list_package_resources()`, either by using a package found via `search_packages()` or `list_packages()`:

```{r list-apartment-building-registration-resources}
apartment_building_registration_package <- search_packages("Apartment Building Registration")

apartment_building_registration_resources <- apartment_building_registration_package %>%
  list_package_resources()

apartment_building_registration_resources
```

or by passing the package's portal URL directly:

```{r list-resources-from-portal}
list_package_resources("https://open.toronto.ca/dataset/apartment-building-registration/")
```

Finally (and most usefully!), you can download the resource (i.e., the actual data) directly into R using `get_resource()`:

```{r get-apartment-building-registration}
library(dplyr)

apartment_building_registration_data <- apartment_building_registration_resources %>%
  filter(name == "Apartment Building Registration Data") %>%
  get_resource()

apartment_building_registration_data
```

The `opendatatoronto` package can currently handle the download of CSV, XLS/XLSX, XML, JSON, SHP, and GeoJSON resources, as well as ZIP resources that contain multiple files. For more information, see the following vignettes:

* [Retrieving multi-sheet XLS/XLSX resources](multisheet_resources.html)
* [Retrieving multi-file ZIP resources](multifile_zip_resources.html)
* [Working with spatial data from the portal](spatial_data.html)
