---
title: "Working with spatial data from the portal"
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

There is a ton of spatial data on the [City of Toronto Open Data Portal](https://open.toronto.ca). Spatial resources are retrieved the same way as all other resources, by using `get_resource()`, and may require the `sf` package.

We can look at the locations of [EarlyON Child and Family Centres](https://open.toronto.ca/dataset/earlyon-child-and-family-centres/) in Toronto. As the portal describes, these centres offer free programs to caregivers and children, providing programs to strengthen relationships, support education, and foster healthy child development. The result of pulling this data in through the package is an `sf` object with WGS84 projection.

```{r setup}
library(opendatatoronto)
library(dplyr)

earlyon_centres <- search_packages("EarlyON Child and Family Centres") %>%
  list_package_resources() %>%
  filter(name == "EarlyON Child and Family Centres Locations - geometry - 4326.zip") %>%
  get_resource()

earlyon_centres
```

If we want to plot this data on a map of Toronto, a shapefile to map the different [neighbourhoods of Toronto](https://open.toronto.ca/dataset/neighbourhoods/) is also available from the portal:

```{r}
neighbourhoods <- list_package_resources("https://open.toronto.ca/dataset/neighbourhoods/") %>%
  filter(name == "Neighbourhoods - 4326.zip") %>%
  get_resource()

neighbourhoods[c("AREA_NA7", "geometry")]
```

Then, we can plot the EarlyON centres along with a map of Toronto:

```{r fig.width = 7}
library(ggplot2)

ggplot() +
  geom_sf(data = neighbourhoods) +
  geom_sf(data = earlyon_centres) +
  theme_void()
```

We may also wish to do something like analyze how many EarlyON centres there are in each neighbourhood. We can count by neighbourhood, using the `sf` package to join the two datasets, then `dplyr` to summarise, and finally `ggiraph` to create an interactive visualization, replacing `geom_sf` with `geom_sf_interactive` and supplying a tooltip:

```{r}
library(sf)
library(dplyr)
library(ggiraph)
library(glue)

earlyon_by_neighbourhood <- neighbourhoods %>%
  st_join(earlyon_centres) %>%
  group_by(neighbourhood = AREA_NA7) %>%
  summarise(n_earlyon = n_distinct(program5, na.rm = TRUE)) %>%
  mutate(tooltip = glue(("{neighbourhood}: {n_earlyon}")))

p <- ggplot() +
  geom_sf_interactive(data = earlyon_by_neighbourhood, aes(fill = n_earlyon, tooltip = tooltip)) +
  theme_void()

girafe(code = print(p))
```

This shows us, for example, that there are 9 EarlyON Centres in West Hill, 5 in Kensington-Chinatown, and 5 in South Riverdale:

```{r}
earlyon_by_neighbourhood %>%
  as_tibble() %>%
  select(neighbourhood, n_earlyon) %>%
  arrange(-n_earlyon) %>%
  head()
```
